---
title: "MAGeCK results"
output:
  html_document:
    df_print: paged
    code_folding: "hide"
---

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = T, warning = F, message = F)

library(tidyverse)
library(data.table)
library(EnhancedVolcano)


theme_set(theme_bw())
cnts = fread("../../resources/crispr/counts/wrangled_raw_counts.tsv")
```

## RRA

Run RRA in 3 contrasts:

-   Day14_Tdneg-vs-Day14_iN\
-   Day14_Tdneg-vs-Day14_Tdpos\
-   Day14_Tdpos-vs-Day14_iN

------------------------------------------------------------------------

### Day14_Tdneg-vs-Day14_iN - look for neuron differentiation activators

For Tdtomato - vs. iN, we look for enriched genes with guides that are lost. TFs with significant reduction in sgRNA reads in the tdTomato- cells indicate that these TFs can potentially positively regulate neuron differentiation.

```{r}
# read in data
tdneg.vs.iN = fread("../../results/crispr/mageckRRA/Day14_Tdneg-vs-Day14_iN/RRA.gene_summary.txt")
```

```{r}
tdneg.vs.iN[, .(id, `neg|score`, `neg|lfc`)] %>%
  column_to_rownames("id") %>% 
  EnhancedVolcano(., rownames(.), x = "neg|lfc", y = "neg|score", 
                  title = NULL, subtitle = NULL, caption = NULL,
                  pCutoff = 1e-3, FCcutoff = 1, max.overlaps = 30,
                  labSize = 3, legendLabSize = 10, legendIconSize = 3,
                  ylim = c(0, 6)
                  ) +
  theme(axis.title = element_text(size = 10),
        axis.text.x = element_text(size = 10), axis.text.y = element_text(size = 10),
        legend.position = "None"
        )
```

### Day14_Tdneg-vs-Day14_Tdpos - look for neuron differentiation activators

For Tdtomato- vs . Tdtomato+ we look for genes with guides that are lost. Similarly to above, TFs with significantly reduced sgRNA reads in the tdTomato- cells vs. tdTomato+ cells may indicate that these TFs can promote neuron differentiation.

```{r}
# read in data
tdneg.vs.tdpos = fread("../../results/crispr/mageckRRA/Day14_Tdneg-vs-Day14_Tdpos/RRA.gene_summary.txt")
```

```{r}
tdneg.vs.tdpos[, .(id, `neg|score`, `neg|lfc`)] %>%
  column_to_rownames("id") %>% 
  EnhancedVolcano(., rownames(.), x = "neg|lfc", y = "neg|score", 
                  title = NULL, subtitle = NULL, caption = NULL,
                  pCutoff = 1e-3, FCcutoff = 1, max.overlaps = 30,
                  labSize = 3, legendLabSize = 10, legendIconSize = 3,
                  ylim = c(0, 6)
                  ) +
  theme(axis.title = element_text(size = 10),
        axis.text.x = element_text(size = 10), axis.text.y = element_text(size = 10),
        legend.position = "None"
        )
```      


### Day14_Tdpos-vs-Day14_iN - look for neuron differentiation inhibitors

In Tdtomato+ vs iN contrast, we look for TFs with significantly reduced sgRNA reads. These indicates that the knock out of these TFs promoted neuron differentiation, suggesting that these TFs are likely inhibitors of neuron differentiation.

<br>

Why do we also use negative screen in tdTomato+ vs. iN? Because this is a CRISPR-cas9 knockout screen, in theory, we should only observe targeted sgRNA reads to decrease.

```{r}
# read in data
tdpos.vs.iN = fread("../../results/crispr/mageckRRA/Day14_Tdpos-vs-Day14_iN/RRA.gene_summary.txt")
```

```{r}
tdpos.vs.iN[, .(id, `neg|score`, `neg|lfc`)] %>%
  column_to_rownames("id") %>% 
  EnhancedVolcano(., rownames(.), x = "neg|lfc", y = "neg|score", 
                  title = NULL, subtitle = NULL, caption = NULL,
                  pCutoff = 1e-3, FCcutoff = 1, max.overlaps = 30,
                  labSize = 3, legendLabSize = 10, legendIconSize = 3,
                  ylim = c(0, 6)
                  ) +
  theme(axis.title = element_text(size = 10),
        axis.text.x = element_text(size = 10), axis.text.y = element_text(size = 10),
        legend.position = "None"
        )
```

## Compare RRA scores


```{r}
inner_join(x = tdneg.vs.iN[, .(id, tdneg_vs_iN = `neg|score`)],
           y = tdneg.vs.tdpos[, .(id, tdneg_vs_tdpos = `neg|score`)], 
           by = "id"
           ) %>% 
  ggplot() + geom_point(aes(tdneg_vs_iN, tdneg_vs_tdpos), alpha = .3, color = "royalblue")
```

```{r}
inner_join(x = tdneg.vs.iN[, .(id, tdneg_vs_iN = `neg|score`)],
           y = tdpos.vs.iN[, .(id, tdpos_vs_iN = `neg|score`)], 
           by = "id"
           ) %>% 
  ggplot() + geom_point(aes(tdneg_vs_iN, tdpos_vs_iN), alpha = .3, color = "firebrick")
```


### Compare lfc

Comparison of log(fold change) shows that indeed, tdneg_vs_iN and tdneg_vs_tdpos is highly correlated. 

```{r}
inner_join(x = tdneg.vs.iN[, .(id, tdneg_vs_iN = `neg|lfc`)],
           y = tdneg.vs.tdpos[, .(id, tdneg_vs_tdpos = `neg|lfc`)], 
           by = "id"
           ) %>% 
  ggplot() + geom_point(aes(tdneg_vs_iN, tdneg_vs_tdpos), alpha = .3, color = "royalblue")
```




```{r}
inner_join(x = tdneg.vs.iN[, .(id, tdneg_vs_iN = `neg|lfc`)],
           y = tdpos.vs.iN[, .(id, tdpos_vs_iN = `neg|lfc`)], 
           by = "id"
           ) %>% 
  ggplot() + geom_point(aes(tdneg_vs_iN, tdpos_vs_iN), alpha = .3, color = "firebrick")
```


## Intersect tdTomato- vs. iN and tdTomato- vs. tdTomato+ top hits

Intersected TFs with `neg|score < 1e-3`. 

```{r}
intersect(tdneg.vs.iN[`neg|score` < 1e-3, id], tdneg.vs.tdpos[`neg|score` < 1e-3, id])
```


Intersects of top 200 TFs

```{r}
intersect(tdneg.vs.iN[`neg|rank` < 201, id], tdneg.vs.tdpos[`neg|rank` < 201, id])
```


top 200 TFs results above, intersecting with our 120 hits. 

```{r}
hits = fread("../../resources/crispr/Hitlist_20191230.csv")
intersect(tdneg.vs.iN[`neg|rank` < 201, id], tdneg.vs.tdpos[`neg|rank` < 201, id]) %>% 
  intersect(hits$gene_name)

```










