---
title: "Respond to ATAC-seq QC comments"
author: "daiuc"
date: "2023-08-14"

---



### Reproducibility

Sample 11 and 12 are day 5, which is drastcally different comparing to other samples. As a result, we notice that sample 11 and 12 are clearly separated from other samples. However, other timepoints may be too close to tell apart when look at all genome wide or the entire set of protein coding genes. 

::: {.panel-tabset}

### Genome wide
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/BigwigSummary/Genome/heatmap.pdf")
```


### Genome wide2
```{r}
#| out-width: "95%"
#| out-height: "500px"

# file produced from analysis/ATAC-QC-replot-corr-heatmap.qmd
knitr::include_graphics("ATAC-QC-bigwig-genomewide-corrplot.pdf")

```



### Genebody
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/BigwigSummary/GeneBody/heatmap.pdf")
```

### Promoter
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/BigwigSummary/Promoter/heatmap.pdf")
```


:::


### Fragment size

In some samples, there is evidence that some samples may be over transposed to a certain extent. However, as demonstrated by the data below, overall size selection bias is kept low across samples. Overall trend indeed shows the highest amount fragment size represent nucleosome free region, and followed mono-nucleosome and di- and tri-nucleosome fragments.

::: {.panel-tabset}

#### ATAC1
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/atac/fragSizeDist/ATAC1.pdf")
```

#### ATAC2
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/atac/fragSizeDist/ATAC2.pdf")
```

#### ATAC3
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/atac/fragSizeDist/ATAC3.pdf")
```

#### ATAC4
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/atac/fragSizeDist/ATAC4.pdf")
```

#### ATAC5
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/atac/fragSizeDist/ATAC5.pdf")
```

#### ATAC6
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/atac/fragSizeDist/ATAC6.pdf")
```

#### ATAC7
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/atac/fragSizeDist/ATAC7.pdf")
```

#### ATAC8
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/atac/fragSizeDist/ATAC8.pdf")
```

#### ATAC9
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/atac/fragSizeDist/ATAC9.pdf")
```

#### ATAC10
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/atac/fragSizeDist/ATAC10.pdf")
```

#### ATAC11
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/atac/fragSizeDist/ATAC11.pdf")
```

#### ATAC12
```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/atac/fragSizeDist/ATAC12.pdf")
```

:::

### Enrichment

We investigated the enrichment of ATAC-seq signals at the promoter region of highly expressed genes with correpsonding time points. We selected a set of 5000 medium to highly expressed genes and plotted ATAC-seq signal strength at the 2kb up and down stream of TSS. The figures below demonstrate the trend that chromatin accessibility is enriched in the promoter regino of these genes, and that we can observe an overall increasing trend of chromatin accessibility as cells further commit to the neuronal developmental path. This increasing trend correlates with our observation in the RNA-seq dataset where significantly more genes are expressed as cells commit towards the neuronal path.

```{r}
#| out-width: "95%"
#| out-height: "500px"

knitr::include_graphics("../code/results/reviews/atac/deeptoolsplots/heatmap_tss_u1k_d1k_top5k_genes.pdf")
```

### Percent of Reads in Peaks

We compared the percent of ATAC-seq reads allocated in peaks. The results show similar observation as we did in the enrichment analysis. As cells are further differentiated towards neurons, more regions of the genome, including protein coding genes / transcription factors, become more accessible. This is reflected in increasing enrichment in the promoter region as we discussed previously, as well as higher percentage of reads in the called peaks regions. Notably, the last two samples, two biological replicates of the final time point, have significantly more reads allocated in the peak regions.

```{r}
#| out-width: "95%"
#| out-height: "600px"

knitr::include_graphics("../code/results/reviews/atac/PRiP.pdf")
```


