---
title: "Compare interaction output files"
author: Chao
date: today
code-fold: true
---

## Introduction

Compare interaction files computed from before (2020) and recomputed now with essentially same code. The purpose is to make sure the results are consistent.

```{r}
#| label: setup
#| include: false

suppressMessages(library(tidyverse))
suppressMessages(library(data.table))
suppressMessages(library(readxl))


```

## Compare interactions (genebody) from current run with run in 2020

```{r}
inter.mx.old = "../code/resources/TFscreen/Interaction_matrix_nofilter_AllTF_20200107.csv"
inter.mx.old = fread(inter.mx.old)

inter.mx.new = "../code/results/reviews/interactions/Interaction_matrix_nofilter_AllTF_genebody_20230111.csv"
inter.mx.new = fread(inter.mx.new)

```

differences in total interaction pairs?
```{r}
interactions = map(
  list(old = inter.mx.old, new = inter.mx.new),
  ~ .x[, .(interaction = paste(regulator_gene, target_gene, sep="_"))] %>% 
    .[, interaction] %>% unique
)

map_int(interactions, length)
```

Intersections of the two is exactly the same number as the length of new interactions. Thus new is a subset of old.
```{r}
reduce(interactions, intersect) %>% length
setdiff(interactions$new, reduce(interactions, intersect)) %>% length
```


What interactions in the old not found in new?
Only about 26 targets comprise genes missed out in the new run.

```{r}
print("Regulators of interactions not found in new:")
setdiff(interactions$old, interactions$new) %>%
  str_split("_", simplify = T) %>% 
  as.data.table %>% 
  pull(V1) %>% unique %>% length

print("Targets of interactions not found in new:")
setdiff(interactions$old, interactions$new) %>%
  str_split("_", simplify = T) %>% 
  as.data.table %>% 
  pull(V2) %>% unique
```

Are the numbers about the same in the new run?

Randomly pick 7 values, show that the difference is almost always 0. Plot shows that
the two runs are mostly exactly the same. 

```{r}
inter.mx.old[, interaction := paste(regulator_gene, target_gene, sep="_")]
inter.mx.new[, interaction := paste(regulator_gene, target_gene, sep="_")]

# common sets
inter.mx = list(
  old = inter.mx.old[interaction %in% reduce(interactions, intersect)][
    order(interaction)],
  new = inter.mx.new[interaction %in% reduce(interactions, intersect)][
    order(interaction)]
)
```

```{r}
# check atac.D4
if (all(inter.mx$old$interaction == inter.mx$new$interaction)) {
  comps = list(
  atac.D4.diff = round(inter.mx$old$atac.D4 - inter.mx$new$atac.D4),
  atac.cnt.D1.diff = round(inter.mx$old$atac.cnt.D1 - inter.mx$new$atac.cnt.D1),
  tar.cnt.ES.diff = round(inter.mx$old$tar.ge.ES - inter.mx$new$tar.ge.ES),
  tar.ge.H16.diff = round(inter.mx$old$tar.ge.H16 - inter.mx$new$tar.ge.H16),
  reg.fdr.D4 = round(inter.mx$old$reg.fdr.D4 - inter.mx$new$reg.fdr.D4, digits = 2),
  tar.l2fc.D1 = round(inter.mx$old$l2fc.D1 - inter.mx$new$tar.l2fc.D1, digits = 2),
  corr = round(inter.mx$old$corr - inter.mx$new$corr, digits = 2)
  ) %>% as.data.table
}

melt(comps, variable.name = "type", value.name = "diff") %>%
  ggplot() + stat_ecdf(aes(diff, color = type))

```





























