---
title: "EncodeChIPSamples"
author: "Chao Dai"
date: "2022-07-21"
output: 
  workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)


library(tidyverse)
library(data.table)
```

## Encode TF ChIP-seq experiments

```{r, include=FALSE}
# encode meta data
encode.meta = "../code/resources/experiment_report_2022_7_19_17h_55m.tsv"
encode.meta = fread(encode.meta)
colnames(encode.meta) = colnames(encode.meta) %>% str_replace_all(" ", "_")
```

All human tissue/cell lines:

```{r}
encode.meta %>% filter(Organism %in% c("Homo sapiens")) %>% 
  select(Organism, Biosample_term_name, Target_gene_symbol, Description) %>% 
  select(Biosample_term_name) %>% unique
```

From the above table, I selected these tissue types as relevant:

    "bipolar neuron", "neural progenitor cell", "neural cell", "astrocyte of the cerebellum",
    "brain", "excitatory neuron", "neural crest cell"

```{r}
neuron_related_tissues = c("bipolar neuron", "neural progenitor cell", "neural cell", 
                           "astrocyte of the cerebellum","brain", "excitatory neuron", 
                           "neural crest cell")
```

ChIP-seq on the following gene products are available:

```{r}
encode.meta %>% 
  select(Organism, Biosample_term_name, Target_gene_symbol, Description) %>% 
  filter(Organism %in% c("Homo sapiens") & Biosample_term_name %in% neuron_related_tissues)
```

```{r}
chip.avail.genes = filter(encode.meta, 
                          Organism %in% c("Homo sapiens") & 
                            Biosample_term_name %in% neuron_related_tissues) %>% 
  pull(Target_gene_symbol) %>% unique

chip.avail.genes
```

## Interaction matrix from Neuron Reporters

```{r}
my.inter = "../code/resources/Interaction_matrix_nofilter_AllTF_20200107.csv"
my.inter = fread(my.inter)
```

There are 456 unique TFs recognized as regulators in our network analyses, shown in table below:

```{r}
select(my.inter, regulator_gene) %>% unique
```

Found only regulators having ChIP-seq data.

```{r}
intersect(my.inter$regulator_gene, chip.avail.genes)
```
